Bayesian multivariate skew-normal finite mixture model for analysis of infant development trajectories

### 2019-06-26 manuscript/manuscript.tex
section 3

\begin{abstract}
In studies of infant motor development, a crucial research goal is to identify latent clusters of infants that experience delayed development, as this is a known risk factor for adverse outcomes later in life. However, there are a number of statistical challenges in modeling infant development: the data are typically skewed, exhibit intermittent missingness, and are highly correlated across the repeated measurements collected during infancy. Using data from the Nurture study, a cohort of over 600 mother-infant pairs followed from pregnancy to 12 months postpartum, we develop a flexible Bayesian latent class model for the analysis infant motor development. Our model has a number of attractive features. First, we adopt the multivariate skew normal distribution with cluster-specific parameters that accommodate the inherent correlation and skewness in the data. Second, we model the cluster membership probabilities using a novel P\'olya-Gamma data-augmentation scheme, thereby improving predictions of the cluster membership allocations. Lastly, we impute missing responses under missing at random assumption by drawing from appropriate conditional skew normal distributions. Bayesian inference is achieved through straightforward Gibbs sampling, and can be carried out in available software such as R.  Through simulation studies, we show that the proposed model yields improved inferences over models that ignore skewness. In addition, our imputation method yields improvements compared to conventional missing data methods, including multiple imputation and complete or available case analysis. When applied to Nurture data, we identified two distinct development clusters: one characterized by delayed “U-shaped” development and a higher percentage of male infants and another characterized by more steady development and a lower percentage of males. The clusters also differed in terms of key demographic variables, such as infant race and maternal pre-pregnancy body mass index. These findings can aid investigators in targeting interventions during this critical early-life developmental window.
\end{abstract}

@@

Mixtures of multivariate non-symmetric distributions such as the multivariate skew-normal (MSN) distribution allow for the nuances of the marginal density to be captured with a more parsimonious set of mixture components. Mixtures of MSN distributions have been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010), however in these models, focus lies primary on marginal density estimation and inference on the mixture components (i.e. clusters) is not discussed. More recently, the mixtures of skew-$t$ factor analysis (MSTFA) model has been proposed for settings in which cluster-specific inference is of primary interest (Lin \textit{et al.} 2018). However, an important feature not included in the MSTFA is the ability to explain individual-level cluster membership as a function of covariates of interest. Additionally, parameter estimation proposed by Lin et al. for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach, namely the ability to incorporate prior information into a model and make posterior probability statements. Our proposed model improves on these previous works by estimating parameters in a Bayesian framework as well as including the ability to fit a multinomial logit regression to cluster membership probabilities using a novel application of data augmentation with the P\'olya Gamma distribution.

@@

A primary goal of the Nurture study is to identify clusters of infants characterized by distinct motor development trajectories. To address this aim, we propose a flexible finite mixture model that accommodates relevant features of the data, such as skewness and dependence among the responses. For $i = 1,...,n$, let $\mathbf{y}_{i}=(y_{i1},\ldots,y_{iJ})^T$ be a $J \times 1$ vector of responses (i.e., Baley scores) for subject $i$ across the $J$ measurement occasions. For the analysis of the Nurture data, we propose a finite mixture model of the form
\begin{equation}
f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ik} f(\mathbf{y}_i|\boldsymbol\theta_k),
\end{equation}
where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,...,K$) and $\pi_{ik}$ is a subject-specific mixing weight representing the probability that subject $i$ belongs to cluster $k$. For now we assume that $K$ is fixed; in Section 4, we discuss model selection strategies for choosing the optimal value of $K$.

To facilitate posterior inference, we introduce a latent cluster indicator variable $z_i$ taking the value $k \in \{1,...,K\}$ with probability $\pi_{ik}$. Conditional on $z_i = k$, we assume $\mathbf{y}_{i}$ is distributed as
\begin{equation}
\mathbf{y}_{i}|(z_i=k) \sim \text{MSN}_J(\boldsymbol\zeta_{ki},\boldsymbol\alpha_k,\boldsymbol\Omega_k), \label{eq:msndens}
\end{equation}
where $\text{MSN}_J(\cdot)$ denotes the $J$-dimensional multivariate skew normal density, $\boldsymbol\zeta_{ki}$ is a $J \times 1$ vector of subject- and cluster-specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of cluster-specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ cluster-specific scale matrix that captures dependence among the $J$ responses for subject $i$. The vector $\boldsymbol\alpha_k$ has components $\alpha_{kj}$, $j = 1,...,J$, that control the skewness of outcome $j$ in cluster $k$. When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $\text{N}_J(\boldsymbol\zeta_k,\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_k$ is a $J \times 1$ mean vector and $\boldsymbol\Omega_k$ is a $J \times J$ covariance matrix.

@@

\subsection{Multinomial Regression for the Cluster Indicators}
To accommodate heterogeneity in the cluster-membership probabilities, we model $\pi_{ik}$ as a function of covariates using a multinomial logit model 
\begin{eqnarray}
\pi_{ik} = \Pr(z_i = k|\mathbf{w}_i) = \frac{\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{h = 1}^K \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}},~ k=1,\ldots,K,
\end{eqnarray}
where $\mathbf{w}_i$ is an $r\times 1$ vector of subject-level covariates, $\boldsymbol\delta_k$ is a $r\times 1$ vector of regression parameters associated with membership in cluster $k$. For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$. Under this model, $z_i|\mathbf{w}_i \sim \text{Multinom}(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$. During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments.

@@

\subsection{Conditional MSN Imputation}
To accommodate missing at random (MAR) responses, we propose a convenient imputation algorithm that can be implemented ``online'' as part of the Gibbs sampler. In Section 6, we discuss extensions to allow for non-ignorable missingness. Suppose $\mathbf{y}_i$ has $q_i\in (1,\ldots,J)$ observed values, denoted $\mathbf{y}^{obs}_i$, and $J-q_i$ intermittent missing values, denoted $\mathbf{y}^{miss}_i$. We can make use of the stochastic representation given in equation (\ref{eq:msnreg}) to impute $\mathbf{y}^{miss}_i$ from its conditional multivariate normal distribution given $(z_i,t_i,\mathbf{y}^{obs}_i)$:
\begin{eqnarray}
\mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& \text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k),~\text{where}\nonumber\\
\boldsymbol\mu^{cond}_{ki}&=& \boldsymbol\mu^{miss}_{ki} + \boldsymbol\Sigma_{12} \boldsymbol\Sigma_{22}^{-1}(\mathbf{y}^{obs}_i - \boldsymbol\mu^{obs}_{ki}) \label{eq:impute}\\
\boldsymbol\Sigma^{cond}_k&=& \boldsymbol\Sigma_{11} - \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}\boldsymbol\Sigma_{21},~\text{and}\nonumber \\
\boldsymbol\Sigma_k &=& \begin{pmatrix} 
	\boldsymbol\Sigma_{k11} & \boldsymbol\Sigma_{k12}\\
	\boldsymbol\Sigma_{k21} & \boldsymbol\Sigma_{k22}
\end{pmatrix}, \ \text{where} \nonumber
\end{eqnarray}
$\boldsymbol\Sigma_{k11}$ is a $J-q_i \times J-q_i$ matrix containing the rows and columns of $\boldsymbol\Sigma_k$ corresponding to inicies of $\mathbf{y}_{i}$ where missingness occurs. Similarly, $\boldsymbol\Sigma_{k12}$ is a $J-q_i \times q_i$ matrix containing the rows of $\boldsymbol\Sigma_k$ that correspond to missing indices of $\mathbf{y}_i$, but columns of $\boldsymbol\Sigma_k$ that correspond to observed indices of $\mathbf{y}_i$. The remaining partitions $\boldsymbol\Sigma_{k21}, \ \text{and} \ \boldsymbol\Sigma_{k22}$ are defined in the same manner. These results follow from conventional multivariate normal theory. An attractive feature of this imputation algorithm is that it avoids multiplicative run-time scaling in $m$, the number of imputations (Gelman \textit{et al.} 2013; Zhou and Reiter, 2010). Our approach also provides more opportunities to explore the missing data parameter space than does multiple imputation, since missing values are drawn at each MCMC iteration, and often in practice $n_{sim} >> m$, where $n_{sim}$ is the number of MCMC iterations (\textbf{find a reference}). In Section 4, we conduct simulation studies to demonstrate that imputing the missing MSN responses improves inferences over complete case analysis.

@@

\subsubsection{Prior Specification}
We adopt a fully Bayesian inferential approach and assign prior distributions to all model parameters. Conveniently, all parameters admit conditionally conjugate priors, which greatly improves posterior computation via a data-augmented Gibbs sampler. For the MSN model component, we adopt a conditionally independent prior structure for $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$, where $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k) = p(\boldsymbol\Sigma_k)p(\boldsymbol\beta_k|\boldsymbol\Sigma_k)$. We choose the normal-inverse-Wishart distribution for $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k)$ by specifying $\boldsymbol\Sigma_k \sim \text{IW}(\mathbf{V}_{0k},\nu_{0k})$ and $\boldsymbol\beta_k|\boldsymbol\Sigma_k \sim \text{N}_{Jp}(\mathbf{b}_k,\mathbf{I}_p \otimes \boldsymbol\Sigma_k)$. We assign the skewness coefficients $\boldsymbol\psi_k$ a conjugate $\text{N}_J(\mathbf{m}_k,\mathbf{P}_k)$ prior. In practice, the updates of $\boldsymbol\beta_k$ and $\boldsymbol\psi_k$ can be combined into one step by defining the $(Jp + J) \times 1$ vector $\boldsymbol\beta^*_k = (\boldsymbol\beta_k,\boldsymbol\psi_k)^T$ for which we assume a $\text{N}_{Jp+J}(\mathbf{b}^*_k,\mathbf{I}_{(p+1)} \otimes \boldsymbol\Sigma_k)$ prior.

To make use of the Matrix Normal representation introduced previously in Section (\textbf{what section}), we define the matrix of regression parameters $\mathbb{B}^*_k$, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k$. We assign $\mathbb{B}^*_k|\boldsymbol\Sigma_k$ a $\text{MatNorm}(\mathbf{B}^*_{0k},\mathbf{I}_{p+1},\boldsymbol\Sigma_k)$, where $\mathbf{B}^*_{0k}$ is a matrix of location parameters such that $\text{vec}(\mathbf{B}^*_{0k}) = \mathbf{b}^*_k$, $\mathbf{I}_{p+1}$ is the identity matrix of dimensions $(p+1)\times(p+1)$, and $\boldsymbol\Sigma_k$ is the model error. This prior structure allows for convinient Matrix Normal updates of the MSN regression parameters as described in Section (\textbf{what section}).

@@

\paragraph{\textbf{Step 1}: Conditional MSN Imputation} The sampler begins with imputing missing values $\mathbf{y}_i^{miss}$ conditional on current simulated values of $z_i = k$ and $t_i$ as well as the associated $\mathbf{y}_i^{obs}$. This is accomplished by drawing $\mathbf{y}_i^{miss}$ from $\text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k)$ as described in equation \ref{eq:impute} for all $i = 1,...,n$. Conditional imputation is completed with construction of a complete outcome vector $\mathbf{y}_i$ by merging the now imputed $\mathbf{y}_i^{miss}$ with $\mathbf{y}_i^{obs}$ with respect for the original order of $\mathbf{y}_i$.

\paragraph{\textbf{Step 2:} Update of MSN Regression Parameters} We begin the update of MSN regression parameters by updating $t_i$, the truncated Normal random effect used in the stochastic representation of $\mathbf{y}_i$ given in equation \ref{eq:msnreg}. For cluster $k$, compute $A_k = (1 + \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k\boldsymbol
\psi_k)^{-1}$ according to current values of $\boldsymbol\psi_k$ and $\boldsymbol\Sigma_k$. Now, for $i = 1,...,n_k$, where $n_k$ is the number of members of cluster $k$ according to the current cluster labeling ${z}_1,...,{z}_n$, compute $a_i = A_k \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k(\mathbf{y}_i - \boldsymbol\zeta_{ik})$. Finally, update $t_i$ from $\text{N}_{[0,\infty)}(a_i, \mathbf{A}_k)$ (\textbf{put derivations in supplement}).  Repeat these updates for $k = 1,...,K$

The remaining MSN regression parameters are updated from their full conditionals as follows. For each cluster $k$ ($k = 1,...,K$) we update $\boldsymbol\Sigma_k$ from $\text{IW}(\nu_k,\mathbf{V}_k)$, where $\nu_k = \nu_{0k} + n_k$ and\begin{eqnarray}
	\mathbf{V}_k &=&  (\mathbf{Y}-\mathbf{X}^*_k \mathbf{B}^*_k)^T (\mathbf{Y}_k - \mathbf{X}^*_k \mathbf{B}^*_k)\nonumber \\
	 &+& (\mathbf{B}_k^* - \mathbf{B}^*_{0k})^T \mathbf{I}_{p+1}(\mathbf{B}^*_k - \mathbf{B}^*_{0k}) + \mathbf{V}_{0k}. \nonumber
\end{eqnarray}We then make use of the Matrix Normal representation introduced in Section 3.2 to draw $\mathbb{B}^*_k$ from $\text{MatNorm}_{p+1,J}(\mathbf{B}_k^*,\mathbf{L}_k^{-1},\boldsymbol\Sigma_k)$, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k = (\beta_{k11},...,\beta_{kJp},\psi_{k1},...,\psi_{kJ})^T$ and
\begin{eqnarray}
	\mathbf{B}^*_k &=& (\mathbf{X}^{*T}_k \mathbf{X}^*_k + \mathbf{I}_{p+1})^{-1} (\mathbf{X}^{*T}_k \mathbf{Y}_k + \mathbf{I}_{p+1} \mathbf{B}_{0k}) \nonumber \\
	\mathbf{L}^*_k &=& \mathbf{X}^{*T}_k \mathbf{X}^*_k + \mathbf{I}_{p+1}. \nonumber
\end{eqnarray}

\paragraph{\textbf{Step 3:}P\'olya--Gamma Data Augmentation for $z_i$} The sampler concludes with updates of the multinomial regression parameters $\boldsymbol\delta_k$, for $k = 1,...,K-1$, followed by updates of each latent cluster indicator $z_i$ $(i=1,\ldots,n)$ from its multinomial logit full conditional. First, we define $U_{ik} = {1}_{z_i = k}$ for $i = 1,...,n$ and $k = 1,...,K-1$. Next, we update $\mathcal{w}_{ik}$ from a $\text{PG}(1,\eta_{ik})$ density, where $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}$, and ${c}_{ik} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$. Next, define $\gamma_{ik} = \frac{u_{ik} - 1/2}{\mathcal{w}_{ik}}$ and let $\boldsymbol\gamma_k = (\gamma_{1k},...,\gamma_{nk})^T$. Finally, for $k = 1,...,K-1$, update $\boldsymbol\delta_k$ from a $\text{N}_r(\mathbf{d}_k,\mathbf{S}_k)$, where $\mathbf{S}_k = (\mathbf{S}_{0k} + \mathbf{W}^T \mathcal{W}_k \mathbf{W})^{-1}$, $\mathcal{W}_k = \text{Diag}(\mathcal{w}_{1k},...,\mathcal{w}_{nk})$,  $\mathbf{W} =  (\mathbf{w}^T_1,...,\mathbf{w}^T_n)^T$, and $\mathbf{d}_k = \mathbf{S}_k (\mathbf{S}_{0k}\mathbf{d}_{0k} + \mathbf{W}^T \mathcal{W}_k \boldsymbol\gamma_k)$.

>>>>

\begin{abstract}
In studies of infant motor development, a crucial research goal is to identify latent clusters of infants that experience delayed development, as this is a known risk factor for adverse outcomes later in life. However, there are a number of statistical challenges in modeling infant development: the data are typically skewed, exhibit intermittent missingness, and are highly correlated across the repeated measurements collected during infancy. Using data from the Nurture study, a cohort of over 600 mother-infant pairs followed from pregnancy to 12 months postpartum, we develop a flexible Bayesian latent class model for the analysis infant motor development. Our model has a number of attractive features. First, we adopt the multivariate skew normal distribution with cluster-specific parameters that accommodate the inherent correlation and skewness in the data. Second, we model the cluster membership probabilities using a novel application of the P\'olya-Gamma data-augmentation scheme, thereby improving predictions of the cluster membership allocations. Lastly, we impute missing responses under missing at random assumption by drawing from appropriate conditional multivariate skew  normal distributions. Bayesian inference is achieved through straightforward Gibbs sampling, and can be implemented in available software such as \texttt{R}.  Through simulation studies, we show that the proposed model yields improved inferences over models that ignore skewness. In addition, our imputation method yields improvements compared to conventional missing data methods, including multiple imputation and complete or available case analysis. When applied to Nurture data, we identified two distinct development clusters: one characterized by delayed “U-shaped” development and a higher percentage of male infants and another characterized by more steady development and a lower percentage of males. The clusters also differed in terms of key demographic variables, such as infant race and maternal pre-pregnancy body mass index. These findings can aid investigators in targeting interventions during this critical early-life developmental window.
\end{abstract}

@@

Mixtures of multivariate non-symmetric distributions such as the multivariate skew-normal (MSN) distribution allow for the nuances of the marginal density to be captured with a more parsimonious set of mixture components. Mixtures of MSN distributions have been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010 and others), however in these models, focus lies primary on marginal density estimation and inference on the mixture components (i.e. clusters) is often not discussed. More recently, the mixtures of skew-$t$ factor analysis (MSTFA) model has been proposed for settings in which cluster-specific inference is of primary interest (Lin \textit{et al.} 2018). However, an important feature not included in the MSTFA is the ability to explain individual-level cluster membership as a function of covariates of interest. Additionally, parameter estimation proposed by Lin et al. for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach, namely the ability to incorporate prior information into a model and make posterior probability statements. Our proposed model improves on these previous works by estimating parameters in a Bayesian framework as well as including the ability to fit a multinomial logit regression to cluster membership probabilities using a novel application of data augmentation with the P\'olya-Gamma distribution.

@@

A primary goal of the Nurture study is to identify clusters of infants characterized by distinct motor development trajectories. To address this aim, we propose a flexible finite mixture model that accommodates relevant features of the data, such as skewness and dependence among the responses. For $i = 1,...,n$, let $\mathbf{y}_{i}=(y_{i1},\ldots,y_{iJ})^T$ be a $J \times 1$ vector of responses (i.e., Baley scores) for subject $i$ across the $J$ measurement occasions. For the analysis of the Nurture data, we propose a finite mixture model of the form
\begin{equation}
f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ik} f(\mathbf{y}_i|\boldsymbol\theta_k),
\end{equation}
where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,...,K$) and $\pi_{ik}$ is a subject-specific mixing weight representing the probability that subject $i$ belongs to cluster $k$. A distinguishing feature of our model is that we treat $\pi_{ik}$ as a gating function (\textbf{cite book}) that depends on a subject's observed data as well as the parameters specific to the cluster to which they belong. For now we assume that $K$ is fixed; in Section 4, we discuss model selection strategies for choosing the optimal value of $K$.

To facilitate posterior inference, we introduce a latent cluster indicator variable $z_i$ taking the value $k \in \{1,...,K\}$ with probability $\pi_{ik}$. Conditional on $z_i = k$, we assume $\mathbf{y}_{i}$ is distributed as
\begin{equation}
\mathbf{y}_{i}|(z_i=k) \sim \text{MSN}_J(\boldsymbol\zeta_{ki},\boldsymbol\alpha_k,\boldsymbol\Omega_k), \label{eq:msndens}
\end{equation}
where $\text{MSN}_J(\cdot)$ denotes the $J$-dimensional multivariate skew normal density, $\boldsymbol\zeta_{ki}$ is a $J \times 1$ vector of subject- and cluster-specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of cluster-specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ cluster-specific scale matrix that captures dependence among the $J$ responses for subject $i$. The vector $\boldsymbol\alpha_k$ has components $\alpha_{kj}$, $j = 1,...,J$, that control the skewness of outcome $j$ in cluster $k$. When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $\text{N}_J(\boldsymbol\zeta_k,\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_k$ is a $J \times 1$ mean vector and $\boldsymbol\Omega_k$ is a $J \times J$ covariance matrix (Azzalini and Dalla Valle, 1996).

@@

\subsection{Multinomial Regression for the Cluster Indicators}
To accommodate heterogeneity in the cluster-membership probabilities, we model $\pi_{ik}$ as a function of covariates using a multinomial logit model 
\begin{eqnarray}
\pi_{ik} = \Pr(z_i = k|\mathbf{w}_i) = \frac{\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{h = 1}^K \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}},~ k=1,\ldots,K,
\end{eqnarray}
where $\mathbf{w}_i$ is an $r\times 1$ vector of subject-level covariates, $\boldsymbol\delta_k$ is a $r\times 1$ vector of regression parameters associated with membership in cluster $k$. For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$. Under this model, $z_i|\mathbf{w}_i \sim \text{Multinom}(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$. During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments as detailed later in section (\textbf{what section}).

@@

\subsection{Conditional MSN Imputation}
To accommodate missing at random (MAR) responses, we propose a convenient imputation algorithm that can be implemented ``online'' as part of the Gibbs sampler. In Section 6, we discuss extensions to allow for non-ignorable missingness. Suppose $\mathbf{y}_i$ has $q_i\in (1,\ldots,J)$ observed values, denoted $\mathbf{y}^{obs}_i$, and $J-q_i$ intermittent missing values, denoted $\mathbf{y}^{miss}_i$. We can make use of the stochastic representation given in equation (\ref{eq:msnreg}) to impute $\mathbf{y}^{miss}_i$ from its conditional multivariate normal distribution given $(z_i,t_i,\mathbf{y}^{obs}_i)$:
\begin{eqnarray}
\mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& \text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k),~\text{where}\nonumber\\
\boldsymbol\mu^{cond}_{ki}&=& \boldsymbol\mu^{miss}_{ki} + \boldsymbol\Sigma_{12} \boldsymbol\Sigma_{22}^{-1}(\mathbf{y}^{obs}_i - \boldsymbol\mu^{obs}_{ki}) \label{eq:impute}\\
\boldsymbol\Sigma^{cond}_k&=& \boldsymbol\Sigma_{11} - \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}\boldsymbol\Sigma_{21},\nonumber \\
\boldsymbol\mu_{ki} &=& \begin{pmatrix} \boldsymbol\mu_{ki}^{miss} \\ \boldsymbol\mu_{ki}^{obs} \end{pmatrix}, \ \text{and} \nonumber \\
\boldsymbol\Sigma_k &=& \begin{pmatrix} 
	\boldsymbol\Sigma_{k11} & \boldsymbol\Sigma_{k12}\\
	\boldsymbol\Sigma_{k21} & \boldsymbol\Sigma_{k22}
\end{pmatrix}, \ \text{where} \nonumber
\end{eqnarray}
$\boldsymbol\Sigma_{k11}$ is a $J-q_i \times J-q_i$ matrix containing the rows and columns of $\boldsymbol\Sigma_k$ corresponding to inicies of $\mathbf{y}_{i}$ where missingness occurs. Similarly, $\boldsymbol\Sigma_{k12}$ is a $J-q_i \times q_i$ matrix containing the rows of $\boldsymbol\Sigma_k$ that correspond to missing indices of $\mathbf{y}_i$, but columns of $\boldsymbol\Sigma_k$ that correspond to observed indices of $\mathbf{y}_i$. The remaining partitions $\boldsymbol\Sigma_{k21}, \ \text{and} \ \boldsymbol\Sigma_{k22}$ are defined in the same manner. Likewise, $\boldsymbol\mu_{ki} = \mathbf{X}_i \boldsymbol\beta_k$ is partitioned into $\boldsymbol\mu_{ki}^{miss}$ and $\boldsymbol\mu_{ki}^{obs}$ with respect to the missing and observed indices of $\mathbf{y}_i$, respectively. These results follow from conventional multivariate normal theory, which we detail in the Web Appendix. An attractive feature of this imputation algorithm is that it avoids multiplicative run-time scaling in $m$, the number of imputations (Gelman \textit{et al.} 2013; Zhou and Reiter, 2010). Our approach also provides more opportunities to explore the missing data parameter space than does multiple imputation, since missing values are drawn at each MCMC iteration, and often in practice $n_{sim} >> m$, where $n_{sim}$ is the number of MCMC iterations (\textbf{find a reference}). In Section 4, we conduct simulation studies to demonstrate that imputing the missing MSN responses improves inferences over available-case analysis.

@@

\subsubsection{Prior Specification}
We adopt a fully Bayesian inferential approach and assign prior distributions to all model parameters. Conveniently, all parameters admit conditionally conjugate priors, which greatly improves posterior computation via a data-augmented Gibbs sampler. For the MSN model component, we adopt a conditionally independent prior structure for $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$, where $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k) = p(\boldsymbol\Sigma_k)p(\boldsymbol\beta_k|\boldsymbol\Sigma_k)$. We choose the normal-inverse-Wishart distribution for $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k)$ by specifying $\boldsymbol\Sigma_k \sim \text{IW}(\mathbf{V}_{0k},\nu_{0k})$ and $\boldsymbol\beta_k|\boldsymbol\Sigma_k \sim \text{N}_{Jp}(\mathbf{b}_k,\mathbf{I}_p \otimes \boldsymbol\Sigma_k)$. We assign the skewness coefficients $\boldsymbol\psi_k$ a conjugate $\text{N}_J(\mathbf{m}_k,\mathbf{P}_k)$ prior. In practice, the updates of $\boldsymbol\beta_k$ and $\boldsymbol\psi_k$ can be combined into one step by defining the $(Jp + J) \times 1$ vector $\boldsymbol\beta^*_k = (\boldsymbol\beta_k^T,\boldsymbol\psi_k^T)^T$ for which we assume a $\text{N}_{Jp+J}(\mathbf{b}^*_k,\mathbf{I}_{(p+1)} \otimes \boldsymbol\Sigma_k)$ prior.

To make use of the Matrix Normal representation introduced previously in Section (\textbf{what section}), we define the matrix of regression parameters $\mathbb{B}^*_k$, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k$. We assign $\mathbb{B}^*_k|\boldsymbol\Sigma_k$ a $\text{MatNorm}(\mathbf{B}^*_{0k},\mathbf{I}_{p+1},\boldsymbol\Sigma_k)$ prior, where $\mathbf{B}^*_{0k}$ is a matrix of location parameters such that $\text{vec}(\mathbf{B}^*_{0k}) = \mathbf{b}^*_k$, $\mathbf{I}_{p+1}$ is the identity matrix of dimensions $(p+1)\times(p+1)$, and $\boldsymbol\Sigma_k$ is the model error. This prior structure allows for convinient Matrix Normal updates of the MSN regression parameters as described in Section (\textbf{what section}).

@@

\paragraph{\textbf{Step 1}: Conditional MSN Imputation} The sampler begins with imputing missing values $\mathbf{y}_i^{miss}$ conditional on current simulated values of $z_i = k$ and $t_i$ as well as the associated $\mathbf{y}_i^{obs}$. This is accomplished by drawing $\mathbf{y}_i^{miss}$ from $\text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k)$ as described in equation \ref{eq:impute} for all $i = 1,...,n$. Conditional imputation concludes with construction of a complete outcome vector $\mathbf{y}_i$ by merging the now imputed $\mathbf{y}_i^{miss}$ with $\mathbf{y}_i^{obs}$ with respect for the original order of $\mathbf{y}_i$.

\paragraph{\textbf{Step 2:} Update of MSN Regression Parameters} We begin the update of MSN regression parameters by updating $t_i$, the truncated Normal random effect used in the stochastic representation of $\mathbf{y}_i$ given in equation \ref{eq:msnreg}. For cluster $k$, compute $A_k = (1 + \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k\boldsymbol
\psi_k)^{-1}$ according to current values of $\boldsymbol\psi_k$ and $\boldsymbol\Sigma_k$. Now, for $i = 1,...,n_k$, where $n_k$ is the number of members of cluster $k$ according to the current cluster labeling ${z}_1,...,{z}_n$, compute $a_i = A_k \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k(\mathbf{y}_i - \boldsymbol\zeta_{ik})$, where $\boldsymbol\zeta_{ik} = X_i \boldsymbol\beta_k$. Finally, update $t_i$ from $\text{N}_{[0,\infty)}(a_i, \mathbf{A}_k)$ (\textbf{put derivations in supplement}).  Repeat these updates for $k = 1,...,K$

The remaining MSN regression parameters are updated from their full conditionals as follows. For each cluster $k$ ($k = 1,...,K$) we update $\boldsymbol\Sigma_k$ from $\text{IW}(\nu_k,\mathbf{V}_k)$, where $\nu_k = \nu_{0k} + n_k$ and\begin{eqnarray}
	\mathbf{V}_k &=&  (\mathbf{Y}-\mathbf{X}^*_k \mathbf{B}^*_k)^T (\mathbf{Y}_k - \mathbf{X}^*_k \mathbf{B}^*_k)\nonumber \\
	 &+& (\mathbf{B}_k^* - \mathbf{B}^*_{0k})^T \mathbf{I}_{p+1}(\mathbf{B}^*_k - \mathbf{B}^*_{0k}) + \mathbf{V}_{0k}. \nonumber
\end{eqnarray}We then make use of the Matrix Normal representation introduced in Section 3.2 to draw $\mathbb{B}^*_k$ from a $\text{MatNorm}_{p+1,J}(\mathbf{B}_k^*,\mathbf{L}_k^{-1},\boldsymbol\Sigma_k)$ density, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k = (\beta_{k11},...,\beta_{kJp},\psi_{k1},...,\psi_{kJ})^T$ and
\begin{eqnarray}
	\mathbf{B}^*_k &=& (\mathbf{X}^{*T}_k \mathbf{X}^*_k + \mathbf{I}_{p+1})^{-1} (\mathbf{X}^{*T}_k \mathbf{Y}_k + \mathbf{I}_{p+1} \mathbf{B}_{0k}) \nonumber \\
	\mathbf{L}^*_k &=& \mathbf{X}^{*T}_k \mathbf{X}^*_k + \mathbf{I}_{p+1}. \nonumber
\end{eqnarray}

\paragraph{\textbf{Step 3:}P\'olya--Gamma Data Augmentation for $z_i$} The sampler concludes with updates of the multinomial regression parameters $\boldsymbol\delta_k$, for $k = 1,...,K-1$, followed by updates of each latent cluster indicator $z_i$ $(i=1,\ldots,n)$ from its multinomial logit full conditional. First, we define $U_{ik} = {1}_{z_i = k}$ for $i = 1,...,n$ and $k = 1,...,K-1$. Next, we update $\mathcal{w}_{ik}$ from a $\text{PG}(1,\eta_{ik})$ density, where $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}$, and ${c}_{ik} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$. Next, define $\gamma_{ik} = \frac{u_{ik} - 1/2}{\mathcal{w}_{ik}}$ and let $\boldsymbol\gamma_k = (\gamma_{1k},...,\gamma_{nk})^T$. Finally, for $k = 1,...,K-1$, update $\boldsymbol\delta_k$ from a $\text{N}_r(\mathbf{d}_k,\mathbf{S}_k)$, where $\mathbf{S}_k = (\mathbf{S}_{0k} + \mathbf{W}^T \mathcal{W}_k \mathbf{W})^{-1}$, $\mathcal{W}_k = \text{Diag}(\mathcal{w}_{1k},...,\mathcal{w}_{nk})$,  $\mathbf{d}_k = \mathbf{S}_k (\mathbf{S}_{0k}\mathbf{d}_{0k} + \mathbf{W}^T \mathcal{W}_k \boldsymbol\gamma_k)$, and $\mathbf{W}$ is the $n \times r$ matrix of multinomial logit regression covariates such that the $i^{th}$ row of $\mathbf{W}$ is $\mathbf{w}_i$.
